---
title: "Tidy tuesday 20-01-14 : the passwords dataset"
output: 
  rmarkdown::html_vignette:
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{td_2020-01-14_Passwords}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 5
)
```


```{r setup, warning=FALSE, message=FALSE}
library(nd.tidytuesday)
library(dplyr)
library(ggplot2)
```


## Data import and inspection

One can obtain the metadata table using `datapasta`'s R package addin and copying the table available at 

```{r metadata}
meta <- data.frame(stringsAsFactors=FALSE,
                   variable = c("rank", "password", "category", "value", "time_unit",
                                "offline_crack_sec", "rank_alt", "strength", "font_size"),
                   class = c("double", "character", "character", "double", "character",
                             "double", "double", "double", "double"),
                   description = c("popularity in their database of released passwords",
                                   "Actual text of the password",
                                   "What category does the password fall in to?", "Time to crack by online guessing",
                                   "Time unit to match with value", "Time to crack offline in seconds",
                                   "Rank 2", "Strength = quality of password where 10 is highest,
                   1 is lowest,
                   please note that these are relative to these generally bad passwords", "Used to create the graphic for KIB")
)
meta <- as_tibble(meta)

pander::pander(meta)
```

the `readr::read_csv` function gives some information on how the data was processed, with here no errors or warnings.

```{r data import}
passwords <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-14/passwords.csv')
```

The passwords dataset has 507 rows * 9 variables. It is not tidy, if we want to use the `value` variable of online guessing we need to add the `time_unit` information to it.

```{r data inspect}
passwords %>% 
  glimpse()
```

Glimpse of the passwords :

```{r data glimpse}
passwords %>% head() %>% pander::pander()
```


## Quality check

```{r data quality}
# --- using base R function
# passwords %>% 
#   qual_fun(.)

# --- tidyverse approach
passwords %>% 
  tidy_qual() %>% 
  pander::pander()
```

The 7 missing values are the in the same rows :

```{r NA}
passwords %>% 
  filter(!complete.cases(passwords))
```

In total those missing values represent less than 2% of the total of data.

```{r naniar vis}
passwords %>% 
  naniar::vis_miss()

passwords %>% 
  visdat::vis_dat()
```


## Cleaning

To fix the types, the character variables are set as factors.

```{r types}
psw_clean <-
  passwords %>% 
  filter(complete.cases(passwords)) %>% 
  mutate_if(is.character, factor) %>% 
  tidyr::unite(data = ., col = "online_duration", value, time_unit, sep = " ", remove = FALSE) %>% 
  mutate(online_duration = lubridate::duration(num = online_duration),
         od_num = as.numeric(online_duration)
  )
```

The value and time unit are united in the `online_time` variable with seconds as unit.

```{r unit time value}
# summary(psw_clean)
log_df <- tibble(
  unit = c("sec", "min", "h", "day", "week", "month", "year", "10years", "100years"),
  logv = log(c( 1, 60, 3600, 24*3600, 7*24*3600, 30*24*3600, 
                365.25*24*3600, 10*365.25*24*3600, 100*365.25*24*3600) ),
  y = 0.65
)

# exp(30)/(20*365.25*24*3600)

psw_clean %>% 
  ggplot(aes(x = log(od_num), y = stat(density))) + 
  geom_histogram() +
  geom_vline(xintercept = log_df$logv, colour = "red") +
  geom_label(data = log_df, aes(x = logv, y = y, label = unit)) + # inherit.aes = FALSE,
  theme_light() + xlab("log(duration)") + ylab("Density") +
  ggtitle("Density historam of the log duration to hack passwords")
```

## Univariate study

There are 10 unique categories and 7 unit of time. Passwords are unique individuals in this dataset.

```{r factor analysis}
psw_clean %>% 
  summarise_if(.tbl = .,.predicate = is.factor, .funs = list("nb_u" = ~length(unique(.))) ) %>% 
  mutate(indic = "nb unique") %>% 
  select(indic, everything())
```


1/3 of the passwords are names, most are actually dictionnary words.

```{r category}
cat_stat <- psw_clean %>%
  group_by(category) %>% 
  summarise(nb = n(),
            pct = nb/nrow(psw_clean)*100) %>% 
  arrange(-nb)

cat_stat %>% 
  ggplot(aes(x = forcats::fct_reorder(.f = category, .x = pct), y = pct)) + 
  geom_col(fill = "#4F62DB") +
  geom_text(aes(y = pct + 2, label = pct), size = 4) +
  xlab("Category of passord") + ylab("Percentage") +
  ggtitle("Repartition of the passwords per categories") +
  theme_light() +
  coord_flip()
```

Nearly 70% of the passwords considered are crackable online in less than a week, 20% are guessable within a day.

```{r time unit}
tu_stat <- psw_clean %>%
  group_by(time_unit) %>% 
  summarise(nb = n(),
            pct = nb/nrow(psw_clean)*100) %>% 
  arrange(-nb)

tu_stat %>% 
  ggplot(aes(x = forcats::fct_reorder(.f = time_unit, .x = pct), y = pct)) + 
  geom_col(fill = "#D91515") +
  geom_text(aes(y = pct + 2, label = pct)) +
  xlab("Time unit to crack a passord") + ylab("Percentage") +
  ggtitle("Repartition of the time unit required to crack passwords") +
  theme_light() +
  coord_flip()
```


We can look at the character type distribution in the passwords.

```{r passwords by character types}
library(stringr)

passwd_type <- psw_clean %>% 
  mutate(password = as.character(password)) %>% 
  mutate("nb_char" = nchar(password),
         num_part = purrr::map_chr( stringr::str_extract_all(string = password, pattern = "[:digit:]"),
                                    function(x) paste(x, collapse = "") ),
         nchar_num = nchar(num_part),
         
         alpha_part = purrr::map_chr( stringr::str_extract_all(string = password, pattern = "[:alpha:]"),
                                      function(x) paste(x, collapse = "") ),
         nchar_alpha = nchar(alpha_part),
         
         punct_part = purrr::map_chr( stringr::str_extract_all(string = password, pattern = "[:punct:]"),
                                      function(x) paste(x, collapse = "") ),
         nchar_punct = nchar(punct_part),
         
         lower_part = purrr::map_chr( stringr::str_extract_all(string = password, pattern = "[:lower:]"),
                                      function(x) paste(x, collapse = "") ),
         nchar_lower = nchar(lower_part),
         
         upper_part = purrr::map_chr( stringr::str_extract_all(string = password, pattern = "[:upper:]"),
                                      function(x) paste(x, collapse = "") ),
         nchar_upper = nchar(upper_part)
  )
```

The data contains passwords with numeric and alpha character. All of alpha character use lower case.

```{r passwords type stats}
passwd_type %>%
  select(dplyr::starts_with("nchar")) %>%
  base_summary() %>% 
  pander::pander()
```

One can suspect a linear relation between the online and offline time of cracking.

```{r passwords viz}
# passwd_type %>% View()
passwd_type %>% # glimpse()
  ggplot(aes(x = od_num, y = offline_crack_sec, colour = strength, size = factor(nb_char)) ) +
  geom_point(alpha = .1)
# facet_grid("category", scales = "free")
```

This is confirmed passing to (log ,log) scaled plot. The formula obatined is 

$$log(offline) = -18.322 + 0.987 \times log(online)$$ 

```{r passwords viz log}
df_min <- passwd_type %>% transmute(x = log(od_num), y = log(offline_crack_sec))
# summary(df_min)
lm_mod <- lm(y ~ x, df_min)
# summary(lm_mod)
df_min$y_hat <- predict.lm(lm_mod, newdata = df_min)

passwd_type %>% # glimpse()
  ggplot(aes(x = log(od_num), y = log(offline_crack_sec), colour = strength, size = nb_char ) ) +
  geom_point(alpha = .1) +
  geom_line(data = df_min, aes(x, y_hat), inherit.aes = FALSE) +
  xlab("Log online time") + ylab("Log offline time") +
  ggtitle("Evolution of the log offline time depending on the log online time") +
  theme_classic()
```



